To achieve ambitious CO$_2$ emission reduction targets quickly, the planning of
energy systems must accommodate societal preferences, such as those regarding
transmission reinforcements or onshore wind parks, and must also acknowledge
uncertainties of technology cost projections among many other uncertainties. To
date, however, many models lean towards only minimising system cost and only
using a single set of cost projections. Here, we address both criticisms in
unison. While taking account of cost uncertainties, we apply multi-objective
optimisation techniques in a fully renewable European electricity system to
explore trade-offs between rising system cost and the deployment of individual
technologies for generating, storing and transporting electricity. We identify
ranges of capacity expansion plans that are cost-efficient and incorporate
uncertainty about future technology cost developments; for instance, we find
that some grid reinforcement and long-term storage alongside large wind
capacities are important to keep costs within 8\% of the least-cost solutions.
We reveal that near the cost-optimum a broad spectrum of technologically diverse
options exist, which allows policymakers to make trade-offs regarding unpopular
infrastructure. For the analysis we carried out more than 50,000 optimisation
runs. To manage many computationally demanding scenario runs efficiently, we
leverage multi-fidelity surrogate modelling techniques using sparse polynomial
chaos expansions and low-discrepancy sampling.
