% summary
In this work, we systematically explore a space of alternatives beyond
least-cost solutions for society and politics to work with subject to uncertain
technology cost projections. We show how narrowly following cost-optimal results
underplays an immense degree of freedom in designing future renewable power
systems. To make our finding that there is no unique path to cost-efficiency
more robust, we account for technology cost uncertainties as one example of the
many unknowns faced in the energy transition, and draw conclusions about the
range of options and cost sensitivities given this uncertainty:

\paragraph{Wide Range of Trade-Offs}
We find that there is a substantial range of options within 8\% of the
least-cost fully renewable electricity system regardless of how cost
developments will unfold. This holds across all technologies individually and
even when considering dependencies between wind and solar, offshore and onshore
wind, as well as hydrogen and battery storage as exemplary flexibility options.

\paragraph{Unlikely Near-Optimal Solutions}
We also carve out parts of the solution space which are unlikely to keep costs
within given cost ranges given the considered range of technology cost futures.
For a fully renewable electricity system, either offshore or onshore wind
capacities of the order of 600 GW along with some long-term storage technology
and transmission network reinforcement appears essential in the scenarios we
analyse.

% \paragraph{Robustness to Cost and Near-Optimal Perturbations}

\paragraph{Key Technology Cost Sensitivities}
We identify onshore wind cost as the apparent main determinant of system cost,
though it can often be substituted with offshore wind for a small additional
cost. This aligns with the finding that the near-optimal feasible space is flat.
Moreover, the deployment of batteries is the most sensitive to its cost, whereas
required levels of transmission expansion are least affected since transmission
cost were not considered to be uncertain. \\

% concluding remark
The robust finding of our study is the consistent investment flexibility in
shaping fully renewable power systems, even without availing of the myriad
flexibility options offered through sector-coupling. This opens the floor to
discussions about social trade-offs and navigating around issues, such as public
opposition towards wind turbines or transmission lines. Rather than modellers
making normative choices about how the energy system should be optimised, by
applying multi-fidelity surrogate modelling techniques and the
modelling-to-generate-alternatives methodology we offer a methodology to present
a wide spectrum of options and trade-offs that are feasible and within a
reasonable cost range, to help society decide how to shape the future of the
energy system.

% ------------------- not used ----------------------------

% Recognising both the parametric and structural uncertainties of our modelling,
% we realise that we--and others--can provide policy-makers with no more, but also no less,
% than a framed canvas and the colors to paint.
% Getting the artwork done is more important than getting it perfect!
% When commissioning studies, it is therefore important that policymakers
% rather ask for a range of plausible choices rather than solutions
% optimised to the last digit.
