% context and motivation

Energy system models have become a pivotal instrument for policy-making to find
cost-efficient system layouts that satisfy ambitious climate change mitigation
targets. But even though they have proliferated in spatial, temporal,
technological and sectoral detail and scope in recent years, least-cost
optimisation models can easily give a false sense of
exactness.\cite{Trutnevyte2016, pye_modelling_2020} Frequently, they present
just a single least-cost solution for a single set of cost assumptions, which
not only neglects uncertainties inherent to technology cost projections which
can have a strong effect on the results of capacity expansion models,
\cite{trondle_trade-offs_2020, yue_review_2018, pye_assessing_2018} but also
hides a wide array of alternative solutions that are equally feasible and only
marginally more expensive.\cite{nearoptimal, lombardi_policy_2020,
sasse_regional_2020}

% relevance of trade-offs

Trade-offs revealed by deviating from least-cost solutions are extremely
attractive for policymakers, because they allow them to make decisions based on
non-economic criteria without affecting the cost-effectiveness of the system.
Knowing that many similarly costly but technologically diverse solutions exist,
helps to accommodate political and social dimensions that are otherwise hard to
quantify; for instance, rising public opposition towards new transmission lines
and onshore wind turbines or an uneven distribution of new infrastructure.
\cite{mccollum_energy_2020,sasse_regional_2020,schlachtberger_cost_2018}

% how to find trade-offs

Techniques like multi-objective optimisation and
modelling-to-generate-alternatives are designed to find such near-optimal
solutions. They have been applied to investment planning models of the
European,\cite{nearoptimal} the Italian, \cite{lombardi_policy_2020} and the
United States power system, \cite{DeCarolis2016} pathways to decarbonise the
power system of the United Kingdom,\cite{Li2017} the European sector-coupled
energy system,\cite{pickeringDiversityOptions2022}, a single-node energy model
of Germany, \cite{nacken_integrated_2019}, global integrated assessment models,
\cite{Price2017} and were combined with a quick hull algorithm to span a
polytope of low-cost solutions for a single set of cost
parameters.\cite{pedersen_modeling_2020}

% trade-offs must be robust to cost uncertainty

However, most of the studies above only use a central cost projection for each
considered technology. Recent decades have shown that many of these projections
contain a high level of uncertainty, particularly for fast-moving technologies
like solar, wind, batteries and hydrogen storage.
\cite{ludererImpactDeclining2021,victoriaSolarPhotovoltaics2021,jaxa-rozenSourcesUncertainty2021}
This uncertainty propagates through the model to strongly affect the optimal and
near-optimal system compositions, thus undermining any analysis of the
trade-offs. Hence, it is crucial that apparent compromises are rigorously tested
for robustness to technology cost uncertainty to raise confidence in conclusions
about viable, cost-effective power system designs. To thoroughly sweep this
uncertainty space, we can avail of previous works on multi-dimensional global
sensitivity analysis techniques in the context of least-cost
optimisation.\cite{trondle_trade-offs_2020, mavromatidis_uncertainty_2018,
pizarro-alonso_uncertainties_2019, fais_impact_2016, usher_value_2015} In this
context, it is also important to note that this methodology can also be used to
address many further uncertainties, for instance, regarding weather variability
between years, demand projections and the level of cross-sectoral integration
that can be realised.

% this work / novelty

In this paper, we bring together near-optimal analysis with global sensitivity
sweeps over uncertainty. This allows us to systematically explore more robust
trade-offs near the cost-optimum of a fully renewable European electricity
system model, PyPSA-Eur,\cite{pypsaeur} and investigate how they are affected by
uncertain technology cost projections. We evaluate compromise solutions between
system cost and technology choices by minimising and maximising the use of on-
and offshore wind, solar photovoltaics, transmission, batteries and hydrogen
storage respectively in order to identify near-optimal alternatives that are no
more than 8\% more expensive than the least-cost option. This discretionary
choice is low enough to argue that such solutions are nearly cost-optimal, and
high enough such that most trade-offs have flattened out at this point. However,
more expensive solutions may still be acceptable to the public, in particular
when they are more widely accepted or if they could be implemented more quickly.
We also show examples of trade-offs for pairs of technologies and a chosen
allowed cost penalty of 6\%; namely between wind and solar, battery and hydrogen
storage, and offshore and onshore wind. For this research, we solve numerous
spatially and temporally explicit long-term investment planning problems that
coordinate generation, transmission and storage investments subject to
multi-period linear optimal power flow constraints. The capacity optimisation is
supplemented with global sensitivity analysis methods to account for a wide
range of cost projections for wind, solar, battery and hydrogen storage
technologies. These cost projections are assumed to be uniformly distributed
based on ranges from the Danish Energy Agency for the year 2050.\cite{DEA}

% computational achievement

To handle the high computational burden incurred by searching for near-optimal
alternatives alongside evaluating many different cost parameter sets, we employ
multi-fidelity surrogate modelling techniques, based on sparse polynomial chaos
expansion, that allow us to merge results from one simpler and another more
detailed model. One covers 37 regions at 4-hourly and the other 128 regions at
2-hourly resolution over a full year (see \nameref{sec:multifidelity} for
details). This approach has been proven very effective in Tröndle et
al.~\cite{trondle_trade-offs_2020}. Heavy parallelisation with high-performance
computing infrastructure allows us to solve more than 50,000 resource-intensive
optimisation problems which, in combination with surrogate modelling, can span a
probabilistic space of near-optimal solutions with respect to cost uncertainties
rather than putting single technology cost futures into the foreground. While
the methodology is general enough to be applied to other uncertainties like
interannual weather variability or demand projections, our application addresses
the probabilistic space of technology cost projections.

% results

In this way, we are able to present alternative solutions beyond least-cost that
have a high chance of involving a limited cost increase, just as we identify
regions that are unlikely to be cost-efficient for realising a fully renewable
electricity system. We derive both ranges of options and technology-specific
capacity expansion ranges that are not affected by cost uncertainty and should
be met to keep the total system cost within a specified range, given the model
setup. Our results show that indeed many such similarly costly but
technologically diverse solutions exist regardless of how technology cost
developments will unfold within the considered ranges.

We approach the presentation of these results in two steps: we first show the
propagation of uncertainties in least-cost solutions, including a global
sensitivity analysis that identifies the most influential cost parameters. We
then gradually expand the uncertainty analysis to the space of nearly
cost-optimal solutions.
